---
title: "Assignment 1"
author: "Yue Ma"
date: "2018/9/24"
output:
  html_document:
    number_sections: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Data Structure

A vector is a list of elements having the same type. It is the fundamental data structure in R. The elements of a vector must all have the same mode or data type. We can have a vector consisting of three character strings (of mode character) or four integer elements (of mode integer), but not a vector with a mix of integer elements and character string elements.

A data frame is a matrix-like data structure in R, with two-dimensional rows and columns where each column may have a different mode or data type. For instance, one column may consist of numbers, and another column might have character strings.

## Vector

```{r}
a <- c(5,6,7,8)
d <- c("orange","red","blue")
print(a)
print(d)
```

## Data.frame

```{r}
name <- c('Alice','Bob','Jack','Jessica','Maria','Derek')
age  <- c(24,36,24,27,24,31)
sex  <- c('F','M','M','F','F','M')
weight <- c(110,120,130,120,140,110)
people <- data.frame(name, age, sex,weight)
people[,c('name','age','sex','weight')]
```

# Style

```{r}
a <- c(5,-2,3,-4,1,NA)
b <- (-a)*(-1)
b[b>0]

num=1
mycondition <- round(runif(1,0,1))
if( mycondition ) {
  num <- num + 1
  print(paste("num =",num,sep=""))
} else{
  print("false")
}

# make an integer sequence from 1 to 50
y <- 1:50

# test if an number is even
even <- y%%2 == 0

# subset by the test above
y = y[even]

x <- matrix(c(23,34,35,6,87,39,21,13,99), nrow=3)
df <- as.data.frame(x)
names(df) <- c("percentage_score_on_reading_test",
               "percentage.score.on.math.test",
               "percentage-score-on-writing-test")

my_mean <- function(x)
{
  sum(x)/length(x)
}
```

# Subsetting a `data.frame`

```{r}
state.x77[c('Connecticut'),c('Population','Life Exp')]
state.x77[7,c(1,4)]
state.x77['Connecticut',c(1,4)]
```

# Manipulating Data

(Note: The code can run successfully and the missing data can be replaced by the average of the non-missing data, however, there is an error that I cannot overcome when I knit it to HTML.)

```{r}
airquality$Solar.R[is.na(airquality$Solar.R)] <- mean(airquality$Solar.R, na.rm = TRUE)
print(airquality)
```

# American National Election Study

```{r}
data("turnout", package = "qss")
help("turnout", package = "qss")
print(turnout)
```

## VAP-based Turnout 

```{r}
VAPtr <- turnout$total/(turnout$VAP + turnout$overseas) * 100
names(VAPtr) <- turnout$year
VAPtr
```

## VEP-based Turnout

```{r}
VEPtr <- turnout$total/turnout$VEP * 100 
names(VEPtr) <- turnout$year
VEPtr
```

## Difference between VEP and VAP-based turnout rates

```{r}
diff <- VEPtr - VAPtr
names(diff) <- turnout$year
diff
```

